IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-07 paper/appendix.tex
Various small improvements, including a revision of extensions.txt (first part).

Here we formally define the syntax and semantics of the extensions
sketched in Section~\ref{sec:extensions}.
In Figure~\ref{fig:labeled-vals} we show how labeled values can be added,
while Figure~\ref{fig:labeled-refs} details labeled references formally.
These changes to the semantics require us to update the erasure function
|erasef il|, which is shown in Figure~\ref{fig:erasure2}.
Note that the soundness proof in Appendix~\ref{sec:appendix}
is done for the fully extended language with labeled values and
labeled references.

\begin{figure}
  %{
  %format .->  = "\overset{\alpha}{\hookrightarrow}"
  %format .-/>  = "\not\overset{\alpha}{\hookrightarrow}"
  \begin{code}
    iv  ::= cdots | Labeled il ie
    ie  ::= cdots | label ie ie | unlabel ie | labelOf ie
    iE  ::= cdots | label iE ie | unlabel iE | labelOf iE
  \end{code}
  \begin{mathpar}
    \inferrule[I-label]
    {|
      il canFlowTo il'
    |}
    {|
      niconf iS id il (label il' ie)
      ->
      niconf iS id il (Labeled il' ie)
    |}
    \and
    \inferrule[I-unlabel]
    {|
      il''=il lub il'
    |}
    {|
      niconf iS id il (unlabel (Labeled il' ie))
      ->
      niconf iS id (il'') ie
    |}
    \and
    \inferrule[I-labelOf]
    {|
    |}
    {|
      niconf iS id il (labelOf (Labeled il' ie))
      ->
      niconf iS id il (il')
    |}
  \end{mathpar}
  \caption{}
  \label{fig:labeled-vals}
  %}
\end{figure}

@@

  \begin{mathpar}
    \inferrule[I-new]
    {
      |il canFlowTo il'| \\
      |fresh ia|\\
      |iS'= bracket iS (mapsto (iall il') iv)|
    }
    {|
      niconf iS id il (new il' iv)
      ->
      niconf iS' id il (iall il')
      |}
    \and
    \inferrule[I-read]
    {}
    {|
      niconf iS id il (read ial')
      ->
      niconf iS id (il lub il') (readState iS (iall il'))
      |}
    \and
    \inferrule[I-write]
    {
      |il canFlowTo il'|\\
      |iS'= bracket iS (mapsto (iall il') iv)|
    }
    {|
      niconf iS id il (write ial' iv)
      ->
      niconf iS' id il (unit)
      |}
    \and
    \inferrule[I-labelOf2]
    {}
    {|
      niconf iS id il (labelOf ial')
      ->
      niconf iS id il (il')
      |}
  \end{mathpar}
  \caption{}
  \label{fig:labeled-refs}
  %}
\end{figure}

\begin{figure}
  \begin{align*}
  &|erase il (iconf iS its) =
  iconf (erase il iS) (fltr (\ it . it = bullet) (map (erasef il) its))|\\
  &|fullconf id il' tS ie| \begin{cases}
  |bullet| & |il' cantFlowTo il| \\
  |fullconf id il' (erase il tS) (erase il ie)| & \text{otherwise}
  \end{cases} \\
  &|erase il (Labeled il' ie)|= \begin{cases}
  |Labeled il' bullet| & |il' cantFlowTo il| \\
  |Labeled il' ie| & \text{otherwise}
  \end{cases} \\
  &|erase il emptyset = emptyset|\\
  &|erase il (bracket iS (mapsto id Q)) =| \begin{cases}
  |erase il iS| & \text{|il' cantFlowTo il|, where |il'| is}\\
  & \text{the label of thread |id|} \\
  |bracket (erase il iS) (mapsto id (erase il Q))| & \text{otherwise}
  \end{cases} \\
  &|erase il (bracket iS (mapsto (iall il') iv)) =| \begin{cases}
  |bracket (erase il iS) (mapsto (iall il') bullet)| & \text{|il' cantFlowTo il|}\\
  |bracket (erase il iS) (mapsto (iall il') (erase il iv))| & \text{otherwise}
  \end{cases} \\
  &|erase il Q = filter il Q|
  \end{align*}
  \caption{Erasure function for the full IFC language, with all extensions.
    In all cases not specified (including target-language constructs),
    |erasef il| is applied homomorphically.
    This definition replaces the one from Figure~\ref{fig:erasure}, which
    is for the IFC language without extensions.}
  \label{fig:erasure2}
\end{figure}

>>>>

Here we formally define the syntax and semantics of the extensions
sketched in Section~\ref{sec:extensions}.
In Figure~\ref{fig:labeled-vals} we show how labeled values can be added,
while Figure~\ref{fig:labeled-refs} details labeled references formally.
For labeled values, we assume that the constructor |Labeled| is not part
of the surface syntax, but rather an internal construct.
These changes to the semantics require us to update the erasure function
|erasef il|, whose full definition is shown in Figure~\ref{fig:erasure2}.
Note that the soundness proof in Appendix~\ref{sec:appendix}
is done for the fully extended language with labeled values and
labeled references.

\begin{figure}
  %{
  %format .->  = "\overset{\alpha}{\hookrightarrow}"
  %format .-/>  = "\not\overset{\alpha}{\hookrightarrow}"
  \begin{code}
    iv  ::= cdots | Labeled il ie
    ie  ::= cdots | label ie ie | unlabel ie | labelOf ie
    iE  ::= cdots | label iE ie | unlabel iE | labelOf iE
  \end{code}
  \begin{mathpar}
    \inferrule[I-label]
    {|
      il canFlowTo il'
    |}
    {|
      niconf iS id il (label il' ie)
      ->
      niconf iS id il (Labeled il' ie)
    |}
    \and
    \inferrule[I-unlabel]
    {|
      il''=il lub il'
    |}
    {|
      niconf iS id il (unlabel (Labeled il' ie))
      ->
      niconf iS id (il'') ie
    |}
    \and
    \inferrule[I-labelOf]
    {|
    |}
    {|
      niconf iS id il (labelOf (Labeled il' ie))
      ->
      niconf iS id il (il')
    |}
  \end{mathpar}
  \caption{Syntax and semantics for labeled values.  These rules are
    understood to be an addition to the existing rules given earlier.}
  \label{fig:labeled-vals}
  %}
\end{figure}

@@

  \begin{mathpar}
    \inferrule[I-new]
    {
      |il canFlowTo il'| \\
      |fresh ia|\\
      |iS'= bracket iS (mapsto (iall il') iv)|
    }
    {|
      niconf iS id il (new il' iv)
      ->
      niconf iS' id il (iall il')
      |}
    \and
    \inferrule[I-read]
    {}
    {|
      niconf iS id il (read ial')
      ->
      niconf iS id (il lub il') (readState iS (iall il'))
      |}
    \and
    \inferrule[I-write]
    {
      |il canFlowTo il'|\\
      |iS'= bracket iS (mapsto (iall il') iv)|
    }
    {|
      niconf iS id il (write ial' iv)
      ->
      niconf iS' id il (unit)
      |}
    \and
    \inferrule[I-labelOf2]
    {}
    {|
      niconf iS id il (labelOf ial')
      ->
      niconf iS id il (il')
      |}
  \end{mathpar}
  \caption{Syntax and semantics for labeled references.  These rules are
    understood to be an addition to the existing rules given earlier.}
  \label{fig:labeled-refs}
  %}
\end{figure}

\begin{figure}
  \begin{align*}
  &|erase il (iconf iS its) =
  iconf (erase il iS) (fltr (\ it . it = bullet) (map (erasef il) its))|\\
  &|fullconf id il' tS ie| \begin{cases}
  |bullet| & |il' cantFlowTo il| \\
  |fullconf id il' (erase il tS) (erase il ie)| & \text{otherwise}
  \end{cases} \\
  &|erase il (Labeled il' ie)|= \begin{cases}
  |Labeled il' bullet| & |il' cantFlowTo il| \\
  |Labeled il' ie| & \text{otherwise}
  \end{cases} \\
  &|erase il emptyset = emptyset|\\
  &|erase il (bracket iS (mapsto id Q)) =| \begin{cases}
  |erase il iS| & \text{|il' cantFlowTo il|, where |il'| is}\\
  & \text{the label of thread |id|} \\
  |bracket (erase il iS) (mapsto id (erase il Q))| & \text{otherwise}
  \end{cases} \\
  &|erase il (bracket iS (mapsto (iall il') iv)) =| \begin{cases}
  |bracket (erase il iS) (mapsto (iall il') bullet)| & \text{|il' cantFlowTo il|}\\
  |bracket (erase il iS) (mapsto (iall il') (erase il iv))| & \text{otherwise}
  \end{cases} \\
  &|erase il Q = filter il Q|
  \end{align*}
  \caption{Erasure function for the full IFC language, with all extensions.
    In all cases that are not specified, including target-language constructs,
    |erasef il| is applied homomorphically
    (e.g., |erase il (setLabel e) = setLabel (erase il e)|).
    This definition replaces the one from Figure~\ref{fig:erasure}, which
    is for the IFC language without extensions.}
  \label{fig:erasure2}
\end{figure}

### 2014-07-07 paper/extensions.tex
Various small improvements, including a revision of extensions.txt (first part).

Different concrete IFC implementations may wish to extend our minimal
system with labeled constructs appropriate for their envisioned
applications.
%
For example, SWAPI provides a labeled version of the XHR object, while
LIO implements a labeled file system, used in server-side web
applications~\cite{hails}.
%
Below we
consider the addition of several labeled constructs to the abstract specification.
%
We remark that that the non-interference proofs must account for the
new constructs, since they amend the IFC calculus semantics (usually
by addition).

\paragraph{Labeled values}
%{
%format Lb = "\textbf{Labeled}"
%format label = "\textbf{label}"
In traditional language-based dynamic IFC systems, a label is
associated with values.
%
Hence, a program that, for example, simply writes labeled messages to
a labeled log can operate on both public and sensitive values.
%
In its simplest form, our coarse grained system requires that the
current label of a task be at least at the level of the sensitive data
to reflect the fact that such data is in scope.

Thus, an extension that supports such fine-grained explicitly
labeled values is essential.
%
To this end, we extend the IFC calculus with explicitly labeled
values, much like those of LIO and
Breeze~\cite{lio, Hritcu:2013:YIB:2497621.2498098}: |tv ::= cdots
inlinesep Lb il tv|.
%
Following LIO, we say that the value |tv| is protected by label |il|,
while the label |il| itself is protected by the task's current label.
%
%
The label of such values can be inspected the task without
requiring the current label to be raised.
%
However, when a task wishes to inspect the protected value |tv|, it
must first raise it label to at least |il| to reflect that it is
incorporating data at such sensitivity level in its scope.
%
When creating labeled values the label |il| must be above
the current label; otherwise it cannot be said that protection has
been transferred from the current label to |il|.
%
(For these rules, we assume that the |Lb| constructor is not part of
the surface syntax and thus cannot be accessed directly to construct
or destruct such values without label checks.)
% 
We omit the reduction rules for labeled values since they closely
follow those presented in~\cite{lio}.
%}

\paragraph{Labeled mutable references/variables/channels}
%
Extending the calculus with other labeled features, such as
references, mutable variables (MVars)~\cite{CH96}, or channels,
follows in a similar way.
%
In these cases, however, we must keep track of the additional state
(as opposed to the labeled value case which requires no additional
bookkeeping).
%
While we can modify the IFC calculus to add such labeled state to each
individual task, adding it to the system state |iS| leads to more
interesting use cases.
%
Specifically, by amending |iS|, as in~\cite{lio,
stefan:addressing-covert}, we can allow threads to use these
constructs to synchronize, or communicate with constructs other than
|send|/|recv| in a safe manner.
%
For example, when extending the calculus with labeled references, |iS|
additionally contains a store that maps addresses to |(il, tv)| pairs
which can be read and written to by different tasks through a labeled
|ref| implementations.
%
Since such labeled constructs have also been considered before in other
concurrent systems, we do not discuss them further.

\paragraph{Privileges}
Decentralized IFC extends IFC with the decentralized label model of
Myers and Liskov~\cite{myers:dlm} to allow for more general
applications, including systems consisting of mutually distrustful
parties.  In a decentralized system, a computation is executed with a
set of \emph{privileges}, which, when exercised, allow the computation
to declassify data (e.g., by lowering the current label).
%
Practical IFC systems
(e.g.,~\cite{Zeldovich:2006, lio,
Hritcu:2013:YIB:2497621.2498098, myers:jif}) rely on privileges to
implement many applications.
%
Since both the LIO and SWAPI implementations already support such features,
we believe that extending our calculus to consider privileges is
straight forward.
%
However, the challenge with such an extension lies in the precise
security guarantees that must be proved.
%
%Addressing this problem is a natural direction for this work.

\paragraph{Clearance}
%
LIO, SWAPI, and Breeze additionally provide a discretionary access
control mechanism---called \emph{clearance}---at the language
level~\cite{Hritcu:2013:YIB:2497621.2498098, lio}; this mechanisms is
used to restrict a computation from accessing data (or communicating
with entities) above a specified label, the clearance.
%
For simplicity, we omitted clearance from our formalism.
%
We only remark that the mechanism generalizes beyond these IFC systems
in a straight forward manner and does not affect our definitions or
proofs in any meaningful way.

>>>>

While the IFC language presented thus far provides the basic
information flow primitives, actual IFC implementations
may wish to extend the minimal system with more specialized
constructs.
%Different concrete IFC implementations may wish to extend our minimal
%system with labeled constructs appropriate for their envisioned
%applications.
%%
For example, SWAPI provides a labeled version of the XHR object, while
LIO implements a labeled file system, used in server-side web
applications~\cite{hails}.
%%
%Below we
%consider the addition of several labeled constructs to the abstract specification.
Below we show how to extend the language with several such constructs,
and highlight some other possible extensions as well.
%
%We remark that that the non-interference proofs must account for the
%new constructs, since they amend the IFC calculus semantics (usually
%by addition).

\paragraph{Labeled values}
%{
%format Lb = "\textbf{Labeled}"
%format label = "\textbf{label}"
In traditional language-based dynamic IFC systems, a label is
associated with values.
%
Hence, a program that, for example, simply writes labeled messages to
a labeled log can operate on both public and sensitive values.
Similarly, a task that receives a sensitive value and forwards it
to another task does not have be be at a sensitive level, if the
value is not inspected.
%
In its simplest form, our coarse grained system requires that the
current label of a task be at least at the level of the sensitive data
to reflect the fact that such data is in scope.

If such fine-grained labeling of values is required, our base IFC
system can be extended with explicitly labeled
values, much like those of LIO and
Breeze~\cite{lio, Hritcu:2013:YIB:2497621.2498098}: |iv ::= cdots
inlinesep Labeled il ie|.
%
Following LIO, we say that the expression |ie| is protected by label |il|,
while the label |il| itself is protected by the task's current label.
%
%
The label of such values can be inspected the task without
requiring the current label to be raised.
%
However, when a task wishes to inspect the protected value |ie|, it
must first raise its label to at least |il| to reflect that it is
incorporating data at such sensitivity level in its scope.
%
When creating labeled values the label |il| must be above
the current label; otherwise it cannot be said that protection has
been transferred from the current label to |il|.
%

Full details with formal reduction rules can be found
in \appendixext{}.
%}

\paragraph{Labeled mutable references/variables/channels}
%
Extending the calculus with other labeled features, such as
references, mutable variables (MVars)~\cite{CH96}, or channels,
can be done in a similar manner.
%
In these cases, however, we must keep track of the additional state
(as opposed to the labeled value case which requires no additional
bookkeeping).
%
%While we can modify the IFC calculus to add such labeled state to each
%individual task, adding it to the system state |iS| leads to more
%interesting use cases.
%
Specifically, by amending |iS|, as in~\cite{lio,
stefan:addressing-covert}, we can allow threads to use these
constructs to synchronize, or communicate with constructs other than
|send|/|recv| in a safe manner.
%
For example, when extending the calculus with labeled references, |iS|
additionally contains a store that maps addresses to a value
and a label
which can be read and written to by different tasks through a labeled
reference implementations.
Again, we omit details here and refer the interested reader
to \appendixext{}, where this extension has been fully worked out.

\paragraph{Privileges}
Decentralized IFC extends IFC with the decentralized label model of
Myers and Liskov~\cite{myers:dlm} to allow for more general
applications, including systems consisting of mutually distrustful
parties.  In a decentralized system, a computation is executed with a
set of \emph{privileges}, which, when exercised, allow the computation
to declassify data (e.g., by lowering the current label).
%
Practical IFC systems
(e.g.,~\cite{Zeldovich:2006, lio,
Hritcu:2013:YIB:2497621.2498098, myers:jif}) rely on privileges to
implement many applications.
%
%Since both the LIO and SWAPI implementations already support such features,
%we believe that extending our calculus to consider privileges is
%straight forward.
%
The challenge with such an extension lies in the precise
security guarantees that must be proved, which to the best of our
knowledge is an open research problem.
%
%Addressing this problem is a natural direction for this work.

\paragraph{Clearance}
%
LIO, SWAPI, and Breeze additionally provide a discretionary access
control mechanism---called \emph{clearance}---at the language
level~\cite{Hritcu:2013:YIB:2497621.2498098, lio}.
This mechanisms is
used to restrict a computation from accessing data (or communicating
with entities) above a specified label, the clearance.
%
For simplicity, we omitted clearance from our formalism.
%
We only remark that the mechanism generalizes beyond these IFC systems
in a straight forward manner and does not affect our definitions or
proofs in any meaningful way.

### 2014-07-07 paper/formal.tex
Various small improvements, including a revision of extensions.txt (first part).

\begin{figure} % fig:erasure
\begin{align*}
  &|erase il (iconf iS its) =
  iconf (erase il iS) (fltr (\ it . it = bullet) (map (erasef il) its))|\\
  &|erase il (fullconf id il' tS ie) =| \begin{cases}
    |bullet| & |il' cantFlowTo il| \\
    |fullconf id il' tS ie| & \text{otherwise}
  \end{cases} \\
  &|erase il emptyset = emptyset|\\
  &|erase il (bracket iS (mapsto id Q)) =| \begin{cases}
    |erase il iS| & \text{|il' cantFlowTo il|, where |il'| is}\\
    & \text{the label of thread |id|} \\
    |bracket (erase il iS) (mapsto id (erase il Q))| & \text{otherwise}
  \end{cases} \\
  &|erase il Q = filter il Q|
\end{align*}
\caption{ Erasure function for tasks, queue maps, message queues, and
configurations.  In all other cases, including target-language constructs,
|erasef il| is applied homomorphically.  \label{fig:erasure} }
\end{figure}

>>>>

\begin{figure} % fig:erasure
\begin{align*}
  &|erase il (iconf iS its) =
  iconf (erase il iS) (fltr (\ it . it = bullet) (map (erasef il) its))|\\
  &|erase il (fullconf id il' tS ie) =| \begin{cases}
    |bullet| & |il' cantFlowTo il| \\
    |fullconf id il' (erase il tS) (erase il ie)| & \text{otherwise}
  \end{cases} \\
  &|erase il emptyset = emptyset|\\
  &|erase il (bracket iS (mapsto id Q)) =| \begin{cases}
    |erase il iS| & \text{|il' cantFlowTo il|, where |il'| is}\\
    & \text{the label of thread |id|} \\
    |bracket (erase il iS) (mapsto id (erase il Q))| & \text{otherwise}
  \end{cases} \\
  &|erase il Q = filter il Q|
\end{align*}
\caption{ Erasure function for tasks, queue maps, message queues, and
configurations.  In all other cases, including target-language constructs,
|erasef il| is applied homomorphically.
Note that |erase il ie| is always equal to |ie| (and similar for |tS|)
in this simple setting.  However,
when the IFC language is extended with more constructs as shown in
Section~\ref{sec:extensions}, then this will no longer be the case.
\label{fig:erasure} }
\end{figure}
